Merging merges, more or less

### 2015-04-08 2015mergingmerges/section/03-Merging.tex
change merge rule names
intro
separate proofs into own section

\begin{tabbing}
MM \= MM \= MMMMMM \= M\kill
$\Psi$ \> $=$  \> @Value@     \> $n$         \\
       \> $~|$ \> @Finished@  \> $n$         \\
       \> $~|$ \> @Closed@    \> $n$         \\
\end{tabbing}

The @merge@ function takes two machines with states $l_1$ and $l_2$, and produces a new machine with state $l_1 \times \{\Psi\} \times l_2 \times \{\Psi\}$.
It starts by creating an empty machine, and inserting states one by one, starting from the initial state as the initial state of each machine, and empty pending sets.
Then, for each output transition of the input machines, states are recursively added.

\begin{tabbing}
MMMMMMM \= MM \= MMMMMM \= M\kill
@merge@ \> $:$ \> $@Machine @l_1 \times @Machine @l_2$ \\
        \> $\to$ \> $@Machine @(l_1 \times \{\Psi\} \times l_2 \times \{\Psi\})$ \\
\\
@merge @$m_1~m_2$ \> $=$ \> $@merge@'~(@init @m_1)~(@init @m_2)~\{\}~@empty@$ \\
\\
$@merge@'$ \> $:$ \> $l_1 \times \{\Psi\} \times l_2 \times \{\Psi\}$               \\
           \> $\times$ \> $@Machine @(l_1 \times \{\Psi\} \times l_2 \times \{\Psi\})$ \\
           \> $\to$ \> $@Machine @(l_1 \times \{\Psi\} \times l_2 \times \{\Psi\})$ \\
\\
MMM \= M \= MMMMMMMMMM \= M \=\kill
$@merge@'~s_1~\psi_1~s_2~\psi_2~m$ \\
 \> $~|$ \> $(s_1, \psi_1, s_2, \psi_2) \in m$  \\
 \> $=$  \> $m$ \\
 \> $~|$ \> $(\gamma,\delta) \in @move@~s_1~\psi_1~s_2~\psi_2$ \\
 \> $=$  \> $@fold@~@merge@'~(m \cup \gamma \cup \delta)~\delta$ \\
\end{tabbing}

@@

\subsection{Theorems we want to prove}

As we do not wish to handle recursive combinators, we also do not need to handle merging cyclic machines.
This is one case that regular dataflow handles but we do not.
The predicate below checks if either of the machines does not use the other's input; if both used the other's input, it would create a cycle.

\begin{tabbing}
MMMMMMM \= MM \= \kill
$@acyclic@~a~b$
\> $:=$
\> $\outputs{a}~\cap~\inputs{b}~=~\emptyset$
\\
\> $~\vee$ \> $\outputs{b}~\cap~\inputs{a}~=~\emptyset$
\end{tabbing}

Two machines cannot be merged if they both provide the same output, or write to the same output channel.

\begin{tabbing}
MMMMMMM \= MM \= \kill
$@distinct@~a~b$
\> $:=$
\> $\outputs{a}~\cap~\outputs{b}~=\emptyset$
\end{tabbing}

\subsubsection{Preservation}
The result of merging two machines should maintain the pull invariants (\S~\ref{s:Machines:Invariants}), provided that the input machines also maintain their invariants.

\begin{tabbing}
MMMMMMM \= MM \= MM \= \kill
$\forall a~b.$
\>
\> $a~@ok@~\wedge~b~@ok@$
\\
\> $\wedge$
\> $@acyclic@~a~b$
\\
\> $\wedge$
\> $@distinct@~a~b$
\\
\> $\implies$
\> $\forall c \in (@merge@~a~b).\ c~@ok@$
\end{tabbing}

\paragraph{Lemma transitions:} for each state $u$ of $@merge@~a~b$ and corresponding states $s$ and $t$ of $a$ and $b$, any output transitions from $u$ either move along $s$'s output transitions or $t$'s output transitions, but not both.
Similarly, the state type of $u$ will either be the type of $s$, the type of $t$, or a @Skip@.

\paragraph{Lemma appears:} for all states $s$ in $a$, if there is a path from $@initial@~a$ to $s$, then:
\begin{itemize}
\item if $s$ is a @Close@, either $s$ or a @Finish@ or a @Close@ from the other machine $b$ will appear in the output ($@merge@~a~b$);
\item if $s$ is an @Out@, @Update@, @If@, $s$ will appear in the output;
\item otherwise, $s$ 
\end{itemize}
ACTUALLY, \textbf{appears} isn't true.

\paragraph{Proof sketch:} each state in $a$ and $b$ has an associated set of $\Psi$.
If we deconstruct each state of $c$, it has a state of $a$, a state of $b$, and a $\Psi$ set for each.

Show that in this $(\sptpC)$, the $\psi$ and $\phi$ is something similar to $s$ and $t$'s $\Psi$ set.

Total $\Psi$ is the union of $\psi$ and $\phi$, except if one set has @Closed@ but the other doesn't, that channel is still open.

Then, look at each rule in @move@, and the preconditions should translate to sufficient preconditions on $\Psi$:

@Skip@, @Update@ and @If@ are trivial.

If both machines are @Done@, since both machines are also @ok@, all their inputs and outputs must be closed..

>>>>

\begin{tabbing}
MMMM \= M \= \kill
@Event@ \> $=$  \> $@Value@~n$ \\
       \> $~|$ \> $@Finished@~n$ \\
       \> $~|$ \> $@Closed@~n$ \\
\end{tabbing}

The @merge@ function takes two machines with states $l_1$ and $l_2$, and produces a new machine with state $l_1 \times \{@Event@\} \times l_2 \times \{@Event@\}$.
It starts by creating an empty machine, and inserting states one by one, starting from the initial state as the initial state of each machine, and empty pending sets.
Then, for each output transition of the input machines, states are recursively added.

\begin{tabbing}
MMMMMMM \= MM \= MMMMMM \= M\kill
@merge@ \> $:$ \> $@Machine @l_1 \times @Machine @l_2$ \\
        \> $\to$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
\\
@merge @$m_1~m_2$ \> $=$ \> $@merge@'~(@init @m_1)~\{\}~(@init @m_2)~\{\}~@empty@$ \\
\\
$@merge@'$ \> $:$ \> $l_1 \times \{@Event@\} \times l_2 \times \{@Event@\}$               \\
           \> $\times$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
           \> $\to$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
\\
MMM \= M \= MMMMMMMMMM \= M \=\kill
$@merge@'~s_1~e_1~s_2~e_2~m$ \\
 \> $~|$ \> $(s_1, e_1, s_2, e_2) \in m$  \\
 \> $=$  \> $m$ \\
 \> $~|$ \> $(\mathit{type},\mathit{transitions}) \in @move@~s_1~e_1~s_2~e_2$ \\
 \> $=$  \> $\bigcup_{t~\in~\mathit{transitions}}(@merge@'~(@snd@~t)~m')$\\
~where  \> $m' = m~\cup~((s_1,e_1,s_2,e_2),~\mathit{type})~\cup~\mathit{transitions}$ \\
\end{tabbing}

@@
